The Occurrence of Earth-Like Planets Around Other Stars

### 2015-03-10 paper/response.txt

For the editor
--------------

As you will see from our response to the referee attached, we have
identified a number of major errors in the referee's discussion of our
work. Additionally, the referee seems unusually hostile to our
modelling choices (e.g. "I could go on with many more effects that are
not adequately modeled, and often not even discussed. But it's not a
referee's job to develop a research plan for them.").  As you will see
in the response, we feel that our choices are well-justified.

Given these issues, could we request to have a second, independent
referee comment on our revised work?  We don't know the procedures for
this, but if appropriate we would be happy for both referees to see
this note to the editor and the attached referee response.

We have made some changes to the discussion in the paper in response
to the current referee's misunderstandings; those are outlined in the
course of the response.

Response
--------

There are several problems with the referee's main points of criticism
of our paper, and also several outright mistakes. Nevertheless, in
some cases we feel that the referee has valid criticisms of our paper.
Below we explain the changes we have made to our manuscript to attempt
to address the referee's criticism and where we feel that criticism is
in error.

The referee writes:

  The entire analysis is predicated on assuming that the intrinsic
  distribution for planets is described by a two-dimensional normal
  distribution and that a piece-wise linear approximation to the
  detection probability that depends only on the estimated transit
  signal to noise. The authors argue that this is a reasonable model,
  based primarily on high signal-to-noise detections of planets much
  larger than Earth and at orbital periods much less than one year.
  However, they they proceed to extrapolate to regions of planet
  radius-orbital period space for which there are very few viable planet
  candidates. Both of their headline results are based on such an
  extrapolation.

The referee is incorrect that our model validation is based on high
signal-to-noise planets at large radii and short periods.  As is
clearly shown in Figure 4 of our manuscript, synthetic data drawn from
our fitted model matches the observations well at:

 * Small planet radius irrespective of period (upper left panel)
 * Peak of the fitted radius distribution (R = 1.25 REarth) (upper
   left panel)
 * Large radius, where the candidates are dominated by false-positives
   (upper left panel)
 * Small periods (lower right panel)
 * Peak of the period distribution (lower right panel)
 * Large period (lower right panel)

It is simply incorrect to say that our model is only validated by
comparison to candidates at much larger than Earth radius and shorter
than one-year periods. The referee *is* correct that there are
relatively few candidates in the P-R plane near 1 year and 1 REarth,
but this is exactly why it is important to validate the model in a
wide range of periods and radii that span the 1 year - 1 REarth region
of interest. We admit to being surprised that such a simple model
reproduced the data so well, but there it is.

The referee writes:

  Further, the manuscript assumes that the apparent size-period
  distribution of false positives can be described by a single
  two-dimensional normal distributions and suggest that this is
  consistent with background eclipsing binary stars being the dominant
  source of false positives among Kepler planet candidates. However,
  previous works have demonstrated that both hierarchical triple stars
  including an eclipsing binary and large planets orbiting the fainter
  member of a binary star are both significant sources of false
  positives. Further, both the rate of false positives and the relative
  abundance of the types of false positives are functions of orbital
  period and planet candidate size. This analysis completely ignores
  these important issues.

Of the four sentences in this paragraph, three contain major errors.
First, our manuscript describes the size-period distribution of false
positives by a linearly-varying function of size and period (c.f. Eq.
(6)), not as a "single two-dimensional normal distribution
[sic]"---perhaps this is a typo? Second, in the third sentence the
referee seems to be complaining that we do not allow for size and
period variations in the background distribution, but it is explicit
in both the text and Figure 1(!) that the distribution of background
contamination we use varies over the size-period plane. Third, in the
final sentence the referee claims that we ignore these issues (related
to contamination), but we do not; for example, we note in the
Conclusion with respect to Petigura et al (2013) and Foreman-Mackey et
al (2014) that "[in contrast to our work] Neither study attempted to
account for contamination from falsely-identified candidate transit
events, controlling this instead through careful choice of threshold."
Similar statements appear throughout the paper. It is our view that
incorporating contamination in the analysis is an important advance
from our paper. In the remaining sentence of this paragraph, the
referee notes that many types of contamination are present in the
data, whose relative importance varies across the size-period plane;
we agree, but note that our contamination model is purely empirical
and agnostic to the source of contamination. While we claim that our
fit is consistent with the amount of contamination from background
eclipsing binaries determined from careful studies, nothing in our
model is based on the detailed properties of this particular source of
contamination.  We have added several sentences throughout the paper
that emphasise this point.  Finally, we have become aware though
helpful comments from Tim Morton that Morton & Johnson (2011) presents
an early measurement of contamination in Kepler, and we now cite that
paper in addition to Fressin et al. (2013) where appropriate.

The referee writes:

  The analysis is based entirely on a publicly available planet
  candidate list, which is heterogeneous, but these issues are not
  addressed. Further, they do not provide enough information about the
  data set used to allow someone else to repeat their calculation.

We state in the introduction that "The Q17 data release describes 2598
“candidate” planetary transit signals identified by the Kepler team
from observations of stars in the “EX” observing program (which are
primarily G-type main-sequence stars similar to our own Sun (Batalha
et al. 2010)), giving the inferred planetary period and radius for
each." and, following Eq. (1) discussing our fiducial SNR, that "In
our analysis, we obtain these quantities from the Kepler Input Catalog
(Batalha et al. 2010; Brown et al. 2011) and the MAST Kepler
archive[web link]." Additionally, we have released all our code for
the analysis at https://github.com/ farr/kepler-selection (linked in
the paper's conclusion). This includes the code that processes the
data from the above-mentioned sources, code that performs the fitting,
and code that produces the plots found in the paper. We are not sure
what additional information the referee would require to repeat our
calculation. We also note that, to our knowledge, no other estimate of
eta-Earth has been performed using this data set (the result from
Petigura et al. 2013 and the following Foreman-Mackey et al. 2014 used
the same lightcurves from Kepler, but a customised pipeline). In some
senses, this data set is heterogeneous (F, G, K stars), but in many
important ways it is not: same instrument (Kepler), same pipeline,
etc. We stand by our claim that this measurement refers to planets
around Sun-like stars, though, unlike Petigura et al. 2013 and
Foreman-Mackey et al. 2014, we did not restrict to G-type stars.

The referee writes:

  Additionally, they do not model the observational uncertainties in
  the estimated planet radii and summarily dismiss a superior
  statistical analysis which both accounts for this effect and
  demonstrates the importance of this effect. The claim that their
  result for eta_Earth is more precise than Foreman-Mackey et al. 2014
  is at best misleading. It reflects a serious misunderstanding of the
  main points from that paper. I am confident that this manuscript's
  estimated uncertainty is smaller, because they ignore important
  effects, extrapolate using a less flexible model, and make additional
  unjustified assumptions.

We do not attempt to model uncertainties in the estimates of period
and radius; Foreman-Mackey et al 2014 do not attempt to account for
contamination. Neither of these choices reflects a mis-understanding
(in fact, Foreman-Mackey et al 2014 devote a paragraph of their
discussion section to "False Positives," noting that their analysis
does not account for them and proposing an extension of their work to
examine their effect), but rather reflects the difficulty of
constructing models that simultaneously reflect both effects. We have
added several sentences to the conclusion section to make this point
clearer. It is true that including period-radius uncertainty might
increase the range of our constraints on eta-Earth and other
parameters, but not by very much; note that we have tested our
sensitivity to uncertainty, stating in Section 2 "In spite of several
candidates with very large uncertainties in measured parameters, we
have found that our fit is essentially unchanged when applied to
synthetic observations with periods and radii re-drawn from the range
of observational uncertainties quoted in the Q17 data release."  We
re-iterate this point now in the conclusion.

The referee writes:

  I could go on with many more effects that are not adequately
  modeled, and often not even discussed. But it's not a referee's job
  to develop a research plan for them.

  In summary, the manuscript presents one basic calculation for an
  interesting, but complex problem which deserves a more detailed
  analysis. I can understand journals accepting one or two early
  analyses that are overly simplistic. But by now the field has plenty
  of simplistic estimates. To make a significant contribution, the field
  needs people to devote the time required to perform a more complete
  analysis that pays attention to both astronomical and statistical
  issues. Such a paper is almost certainly longer than appropriate for
  ApJL. But there is certainly room for at least several more analyses
  that build on each other to develop a progressively more complete
  model to understand the frequency and distribution of small planets
  around other stars.

We disagree with the referee that our calculation is "basic" or
"overly simplistic." If it were basic to incorporate contamination or
empirically estimate selection effects then I am sure that previous
papers would have incorporated elements of our analysis. Instead, we
feel that our novel statistical methods for approaching this problem,
applied to a previously-unexamined data set provide a useful
independent estimate of the size-period distribution of exoplanets
that confirms some aspects of previous analyses and calls others into
question. It is true that our model of the distribution and selection
functions is simple---but it fits the data extremely well. Should we
complicate it in spite of this?  The referee seems unusually hostile
to our modelling choices; given the large number of errors in the
referee's statements about our paper, we would like to request a
second referee's opinion on our paper.

### 2015-03-10 paper/kepler-selection.tex
Cleanup of references and re-format properly for ApJL.
emulateapj apparently doesn't include the proper bibliography style files, so
added hapj.
Changed Ilya's e-mail.
Replaced non-printed \and with \& in author list; testing git and Latex installation on new laptop. :)
Proposed changes for response to referee report.

The quantity $\etaearth$, the number density of planets per star per
logarithmic planetary radius per logarithmic orbital period at one
Earth radius and one year period, describes the occurrence of
Earth-like extrasolar planets.  Measurement of $\etaearth$ is
complicated by the difficulty of detecting Earth-like planets in
Earth-like orbits about Sun-like stars.  Here we present a measurement
of $\etaearth$ from a parameterised forward model of the (correlated)
period-radius distribution and the observational selection function in
the most recent (Q17) data release from the Kepler satellite
\citep{Borucki2010,Borucki2011,Batalha2013}.  Our data set comprises
181,568 systems observed under the Kepler exoplanet observing program
(mostly G-type stars on the main sequence \citep{Batalha2010}),
producing 2598 planetary candidates.  We parameterise the distribution
of planetary periods and radii using a single, correlated Gaussian
component; treat selection effects using a parameterised transit
detection probability based on the measured noise level and stellar
properties in the Kepler catalog; and include an
empirically-parameterised, independent component in the period-radius
distribution to represent false-positive planet detections.  Using our
model we can simultaneously estimate $\etaearth$, place constraints on
the planet period-radius distribution function, and determine the
degree of contamination by false-positive candidate identifications.
We find $\etaearth = \earange$ (90\% CL).  We conclude that each star
hosts $\rplrange$ planets with $P \lesssim 3 \mathrm{yr}$ and $R
\gtrsim 0.2 \REarth$, that the peak of the planet radius distribution
lies at $\Rpeak = \rpeakrange \REarth$, and that $\ln P$ and $\ln R$
are correlated with correlation coefficient $r = \corrcoeffrange$ (all
90\% CL).  Our empirical model for false-positive contamination is
consistent with the dominant source being background eclipsing binary
stars \citep{Fressin2013}, with $\fposrange$ (90\% CL) of the
candidates being false-positives.  The distribution of planets we
infer is consistent with a highly-stochastic planet formation process
producing many correlated, fractional changes in planet sizes and
orbits.  Our approach of determining both the intrinsic distribution
of objects and selection effects empirically from survey data is
generally applicable.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{bg}
  \caption{\label{fig:det-bg} \textbf{Inferred detection probability
      and density of background contamination.} (Top) The inferred
    detection probability versus signal-to-noise ratio (see
    Eq.\ \eqref{eq:pdetect}) from our parameterised model of selection
    effects.  The solid line is the posterior median detection
    probability and the shading gives the 90\% credible posterior
    interval.  Our inferred detection probability is in rough
    agreement with the measurements of detection efficiency in
    \citet{Borucki2011,Batalha2013}.  (Bottom) The number density of
    false-positive candidate signals, $d\Nbg/d\ln P \, d\ln R$ (see
    Eq.\ \eqref{eq:background-rate}).  The density is highest at small
    candidate period and large radius, consistent with the dominant
    source of contamination being background eclipsing
    binaries \citep{Fressin2013}.  Overall, our model finds
    $\fposrange$ of the candidates are false-positive background
    signals, consistent with the analysis in \citet{Fressin2013}.}
\end{figure}

@@

In addition to true planetary signals, we model a false-positive
background of planet candidates empirically, assuming they populate
the candidate $P$-$R$ plane with a number density that has a linear
gradient across a rectangular region in the $\ln P$-$\ln R$ plane:
\begin{multline}
  \label{eq:background-rate}
  \frac{d\Nbg}{d \ln P \, d \ln R} = \frac{\Nbg}{\Delta \ln P \Delta
    \ln R} \\ \times \left( 1 + \vec{\gamma} \cdot \left[ \ln P - \ln P_\mathrm{mid} , \ln
    R - \ln R_\mathrm{mid} \right] \right),
\end{multline}
where $\Delta \ln P = \ln P_\mathrm{max} - \ln P_\mathrm{min}$, $\ln
P_\mathrm{mid} = 1/2\left(\ln P_\mathrm{max} - \ln P_\mathrm{min}
\right)$, $\Delta \ln R = \ln R_\mathrm{max} - \ln R_\mathrm{min}$,
$\ln R_\mathrm{mid} = 1/2\left(\ln R_\mathrm{max} - \ln R_\mathrm{min}
\right)$.  $\Nbg$, the expected number of background false-positive
events; $P_\mathrm{max}$, $P_\mathrm{min}$, $R_\mathrm{max}$, and
$R_\mathrm{min}$, the boundaries in the $P$-$R$ plane within which
background events appear; and $\gamma$, the gradient in the number
density of background events, are parameters of our model.  This is a
purely empirical model for the background contamination, but is
reasonable if the chief contaminant is background eclipsing binaries
\citep{Fressin2013,Duquennoy1991}.  The posterior on the background
number density in the $P$-$R$ plane appears in
Figure\ \ref{fig:det-bg}.

@@

Our model allows us to produce a posterior on the distribution of
planets in the period-radius plane, and the probability that any given
planetary candidate is a planet instead of a background contaminant;
these posteriors appear in Fig. \ref{fig:foreground-dist}.  Our model
finds that the false-positive rate in the candidate data set is
$\fposrange$ (90\% CL), consistent with previous work
\citep{Fressin2013} estimating the contamination in the Kepler
candidate set.  Our model has the peak of the planet period-radius
distribution at $\Rpeak = \rpeakrange \REarth$, $P_\mathrm{peak} =
\ppeakrange \mathrm{yr}$, and the distribution of planetary radii and
periods is correlated, with correlation coefficient $r =
\corrcoeffrange$ (all at 90\% CL).

\begin{figure}
  \includegraphics[width=\columnwidth]{foreground-dist}
  \caption{\label{fig:foreground-dist} \textbf{The inferred planet
      period--radius distribution accounting for selection effects and
      false-positives.}  (Upper Left) The planet number density per
    logarithmic planet radius.  The density peaks at $\Rpeak =
    \rpeakrange \REarth$ (90\% CL).  (Upper Right) The planet number
    density in the period--radius plane.  The inferred correlation
    coefficient between $\ln P$ and $\ln R$ is $r = \corrcoeffrange$.
    (Lower Left) Scatter plot of the radius and period of the Kepler
    planet candidates.  Color indicates the posterior false-positive
    probability for each candidate.  Overall, the model prefers a
    false-positive rate of $\fposrange$ (90\% CL).  The primary
    contaminant is probably background eclipsing binaries; our
    contamination rate is consistent with previous work
    \citep{Fressin2013}. (Lower Right) The planet number density per
    logarithmic planet period.  The density peaks at $P = \ppeakrange
    \mathrm{yr}$ (90\% CL). }
\end{figure}

@@

Previous estimates
\citep{Catanzarite2011,Traub2012,Dong2013,Petigura2013,Foreman-Mackey2014}
place $1\% \lesssim \etaearth \lesssim 34\%$.  These works dealt with
the problem of selection effects in the sample by either analysing a
region of the period-radius parameter space where observations are
complete and extrapolating to $R = \REarth$ and $P = 1 \mathrm{yr}$
\citep{Catanzarite2011,Traub2012}, applying a binned analysis
incorporating survey incompleteness in the period-radius plane
\citep{Dong2013,Petigura2013} or analysing the results of a customised
planet detection pipeline on a subset of the Kepler observations
\citep{Petigura2013,Foreman-Mackey2014}.  The methods and analysed
data sets of \citet{Petigura2013,Foreman-Mackey2014} are most
comparable to ours.  These studies used the same data set, produced
\citep{Petigura2013} from a subset of the available Kepler data and a
customised pipeline to search for transit signals.  They both
accounted for selection effects by measuring the recoverability of
synthetic transit signals injected into their data, in contrast to our
approach of empirically determining them from the observed data.
Neither study attempted to account for contamination from
falsely-identified candidate transit events, controlling this instead
through careful choice of threshold.  Both studies used a more
flexible model for the intrinsic distribution of planets than ours.
Our result for $\etaearth$ is consistent with, but more precise than,
\citet{Foreman-Mackey2014} and (somewhat) inconsistent with
\citet{Petigura2013}.

@@

The code implementing this analysis is available under an open-source
``MIT'' license at \url{https://github.com/farr/kepler-selection}.
This work was supported by the Science and Technology Facilities
Council.  Computations in this work were performed on the University
of Birmingham's BlueBEAR cluster.  Some of the data presented in this
paper were obtained from the Mikulski Archive for Space Telescopes
(MAST). STScI is operated by the Association of Universities for
Research in Astronomy, Inc., under NASA contract NAS5-26555. Support
for MAST for non-HST data is provided by the NASA Office of Space
Science via grant NNX13AC07G and by other grants and contracts.  This
paper includes data collected by the Kepler mission. Funding for the
Kepler mission is provided by the NASA Science Mission directorate.

\bibliography{kepler-selection}

>>>>

The quantity $\etaearth$, the number density of planets per star per
logarithmic planetary radius per logarithmic orbital period at one
Earth radius and one year period, describes the occurrence of
Earth-like extrasolar planets.  Measurement of $\etaearth$ is
complicated by the difficulty of detecting Earth-like planets in
Earth-like orbits about Sun-like stars.  Here we present a measurement
of $\etaearth$ from a parameterised forward model of the (correlated)
period-radius distribution and the observational selection function in
the most recent (Q17) data release from the Kepler satellite
\citep{Borucki2010,Borucki2011,Batalha2013}.  Our data set comprises
181,568 systems observed under the Kepler exoplanet observing program
(mostly G-type stars on the main sequence \citep{Batalha2010}),
producing 2598 planetary candidates.  We parameterise the distribution
of planetary periods and radii using a single, correlated Gaussian
component; treat selection effects using a parameterised transit
detection probability based on the measured noise level and stellar
properties in the Kepler catalog; and include an
empirically-parameterised, independent component in the period-radius
distribution to represent false-positive planet detections.  Using our
model we can simultaneously estimate $\etaearth$, place constraints on
the planet period-radius distribution function, and determine the
degree of contamination by false-positive candidate identifications.
We find $\etaearth = \earange$ (90\% CL).  We conclude that each star
hosts $\rplrange$ planets with $P \lesssim 3 \mathrm{yr}$ and
$R \gtrsim 0.2 \REarth$, that the peak of the planet radius
distribution lies at $\Rpeak = \rpeakrange \REarth$, and that $\ln P$
and $\ln R$ are correlated with correlation coefficient
$r = \corrcoeffrange$ (all 90\% CL).  Our empirical model for
false-positive contamination does not depend on the detailed
properties of the source of contamination, but is consistent with the
dominant source being background eclipsing binary stars
\citep{Morton2011,Fressin2013}, with $\fposrange$ (90\% CL) of the
candidates being false-positives.  The distribution of planets we
infer is consistent with a highly-stochastic planet formation process
producing many correlated, fractional changes in planet sizes and
orbits.  Our approach of determining both the intrinsic distribution
of objects and selection effects empirically from survey data is
generally applicable.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{bg}
  \caption{\label{fig:det-bg} \textbf{Inferred detection probability
      and density of background contamination.} (Top) The inferred
    detection probability versus signal-to-noise ratio (see Eq.\
    \eqref{eq:pdetect}) from our parameterised model of selection
    effects.  The solid line is the posterior median detection
    probability and the shading gives the 90\% credible posterior
    interval.  Our inferred detection probability is in rough
    agreement with the measurements of detection efficiency in
    \citet{Borucki2011,Batalha2013}.  (Bottom) The number density of
    false-positive candidate signals, $d\Nbg/d\ln P \, d\ln R$ (see
    Eq.\ \eqref{eq:background-rate}).  The density is highest at small
    candidate period and large radius, consistent with the dominant
    source of contamination being background eclipsing binaries
    \citep{Morton2011,Fressin2013} (but note that our model does not
    depend on the details of the sources contamination).  Overall, our
    model finds $\fposrange$ of the candidates are false-positive
    background signals, consistent with the analysis in
    \citet{Morton2011,Fressin2013}.}
\end{figure}

@@

In addition to true planetary signals, we model a false-positive
background of planet candidates empirically, assuming they populate
the candidate $P$-$R$ plane with a number density that has a linear
gradient across a rectangular region in the $\ln P$-$\ln R$ plane:
\begin{multline}
  \label{eq:background-rate}
  \frac{d\Nbg}{d \ln P \, d \ln R} = \frac{\Nbg}{\Delta \ln P \Delta
    \ln R} \\ \times \left( 1 + \vec{\gamma} \cdot \left[ \ln P - \ln P_\mathrm{mid} , \ln
    R - \ln R_\mathrm{mid} \right] \right),
\end{multline}
where $\Delta \ln P = \ln P_\mathrm{max} - \ln P_\mathrm{min}$,
$\ln P_\mathrm{mid} = 1/2\left(\ln P_\mathrm{max} - \ln P_\mathrm{min}
\right)$,
$\Delta \ln R = \ln R_\mathrm{max} - \ln R_\mathrm{min}$,
$\ln R_\mathrm{mid} = 1/2\left(\ln R_\mathrm{max} - \ln R_\mathrm{min}
\right)$.
$\Nbg$, the expected number of background false-positive events;
$P_\mathrm{max}$, $P_\mathrm{min}$, $R_\mathrm{max}$, and
$R_\mathrm{min}$, the boundaries in the $P$-$R$ plane within which
background events appear; and $\gamma$, the gradient in the number
density of background events, are parameters of our model.  This is a
purely empirical model for the background contamination, not dependent
on any properties of the source of contamination, but is reasonable if
the chief contaminant is background eclipsing binaries
\citep{Morton2011,Fressin2013,Duquennoy1991}.  The posterior on the
background number density in the $P$-$R$ plane appears in Figure\
\ref{fig:det-bg}.

@@

Our model allows us to produce a posterior on the distribution of
planets in the period-radius plane, and the probability that any given
planetary candidate is a planet instead of a background contaminant;
these posteriors appear in Fig. \ref{fig:foreground-dist}.  Our model
finds that the false-positive rate in the candidate data set is
$\fposrange$ (90\% CL), consistent with previous work
\citep{Morton2011,Fressin2013} estimating the contamination in the
Kepler candidate set.  Our model has the peak of the planet
period-radius distribution at $\Rpeak = \rpeakrange \REarth$,
$P_\mathrm{peak} = \ppeakrange \mathrm{yr}$, and the distribution of
planetary radii and periods is correlated, with correlation
coefficient $r = \corrcoeffrange$ (all at 90\% CL).

\begin{figure}
  \includegraphics[width=\columnwidth]{foreground-dist}
  \caption{\label{fig:foreground-dist} \textbf{The inferred planet
      period--radius distribution accounting for selection effects and
      false-positives.}  (Upper Left) The planet number density per
    logarithmic planet radius.  The density peaks at $\Rpeak =
    \rpeakrange \REarth$ (90\% CL).  (Upper Right) The planet number
    density in the period--radius plane.  The inferred correlation
    coefficient between $\ln P$ and $\ln R$ is $r = \corrcoeffrange$.
    (Lower Left) Scatter plot of the radius and period of the Kepler
    planet candidates.  Color indicates the posterior false-positive
    probability for each candidate.  Overall, the model prefers a
    false-positive rate of $\fposrange$ (90\% CL).  The primary
    contaminant is probably background eclipsing binaries; our
    contamination rate is consistent with previous work
    \citep{Morton2011,Fressin2013}. (Lower Right) The planet number
    density per logarithmic planet period.  The density peaks at $P =
    \ppeakrange \mathrm{yr}$ (90\% CL). }
\end{figure}

@@

Previous estimates
\citep{Catanzarite2011,Traub2012,Dong2013,Petigura2013,Foreman-Mackey2014}
place $1\% \lesssim \etaearth \lesssim 34\%$.  These works dealt with
the problem of selection effects in the sample by either analysing a
region of the period-radius parameter space where observations are
complete and extrapolating to $R = \REarth$ and $P = 1 \mathrm{yr}$
\citep{Catanzarite2011,Traub2012}, applying a binned analysis
incorporating survey incompleteness in the period-radius plane
\citep{Dong2013,Petigura2013} or analysing the results of a customised
planet detection pipeline on a subset of the Kepler observations
\citep{Petigura2013,Foreman-Mackey2014}.  The methods and analysed
data sets of \citet{Petigura2013,Foreman-Mackey2014} are most
comparable to ours.  These studies used the same data set, produced
\citep{Petigura2013} from a subset of the available Kepler data and a
customised pipeline to search for transit signals.  The official
Kepler pipeline used to produce the candidate list used in this work
is an independent code that searches the same lightcurves for planets
using different algorithms. Both \citet{Petigura2013} and
\citet{Foreman-Mackey2014} accounted for selection effects by
measuring the recoverability of synthetic transit signals injected
into their data, in contrast to our approach of empirically
determining them from the observed data.  Neither study attempted to
account for contamination from falsely-identified candidate transit
events as we do, controlling this instead through careful choice of
threshold.  Both studies used a more flexible model for the intrinsic
distribution of planets than ours, and \citet{Foreman-Mackey2014}
accounted for the observational uncertainty in periods and radii
explicitly in their model.  \citet{Foreman-Mackey2014} found that the
discrepancy between their results and those of \citet{Petigura2013}
was likely due to the un-modelled uncertainty in period and radius in
the latter work.  Our result for $\etaearth$ is consistent with, but
more precise than, \citet{Foreman-Mackey2014} and inconsistent with
\citet{Petigura2013}.  We have checked that our results remain largely
unchanged when re-analysing synthetic data sets that contain periods
and radii drawn from the range of observational uncertainties quoted
by the Kepler pipeline.  Combining the methods of
\citet{Foreman-Mackey2014} for treating measurement uncertainty with
the methods of this work for dealing with contamination is left for
future work.

@@

We thank Timothy D. Morton for helpful comments on this manuscript.
The code implementing this analysis is available under an open-source
``MIT'' license at \url{https://github.com/farr/kepler-selection}.
This work was supported by the Science and Technology Facilities
Council.  Computations in this work were performed on the University
of Birmingham's BlueBEAR cluster.  Some of the data presented in this
paper were obtained from the Mikulski Archive for Space Telescopes
(MAST). STScI is operated by the Association of Universities for
Research in Astronomy, Inc., under NASA contract NAS5-26555. Support
for MAST for non-HST data is provided by the NASA Office of Space
Science via grant NNX13AC07G and by other grants and contracts.  This
paper includes data collected by the Kepler mission. Funding for the
Kepler mission is provided by the NASA Science Mission directorate.

\bibliographystyle{hapj}
\bibliography{kepler-selection}
